Optimal Symmetric Multimodal Templates and Concatenated Random Forests for Supervised Brain Tumor Segmentation (Simplified) with \textit{ANTsR}

### 2014-04-29 paper.tex
ENH: More edits.

We use random forests as supervised learners.  Given ground truth,
voxel-wise $n$-class labels for a specific image, we can vectorize the
label image to produce $v_L$, the $p$-length vector of class labels
where $p$ is the number of voxels in the given brain of interest.
We also assume a matrix of features for the given brain, which we
denote $\bm{F}$ where $\bm{F}$ has dimensionality of $p$ by
$k$, where each of the $k$ columns corresponds a potentially valuable
feature vector.  Given this representation, the random forest model
can be trained, in \textit{R} notation, to predict the labels via
$v_L \sim  \bm{F}$.
Our segmentation protocol involves training and application
of two RF models in succession 
(see Figure \ref{fig:pipeline}) to which we refer as ``stages.''  The basic idea is that
we generate a set of feature images used as input to the
first RF model (or first stage) which produces a voxelwise 
probabilistic tissue estimate.  More precisely, in the RF framework, each prediction sample (i.e. the feature vector at each voxel),
is propagated through each tree of the ensemble where it is labeled as belonging to a specific
class.  
These ``votes'' are converted to voxelwise probabilistic
estimates for each class via standard mechanisms used in random forest
models, e.g. \cite{liaw2002}.
The seven labels of brain/tumor tissue type denote:
\begin{itemize}
\item cerebrospinal fluid,
\item grey matter,
\item white matter,
\item edema, 
\item non-enhancing tumor (including low-grade tumor center), 
\item enhancing tumor (excluding necrotic center), and 
\item abnormal necrotic center or necrocyst in high-grade gliomas.
\end{itemize}
We then use these output tissue map estimates as spatial 
priors for generating a second set of geometrically refined image features.  These
are used as input for a second RF model application (or
second stage),
the output of which constitute the final tumor segmentation estimate.
Note that some feature images are included directly and
  without modification in both stages such as
the asymmetry features described in the previous section.

>>>>

We use random forests as supervised learners.  Given ground truth,
voxel-wise $n$-class labels for a specific image, we can vectorize the
label image to produce $v_L$, the $p$-length vector of class labels
where $p$ is the number of voxels in the given brain of interest.
We also assume a matrix of features for the given brain, which we
denote $\bm{F}$ where $\bm{F}$ has dimensionality of $p$ by
$k$, where each of the $k$ columns corresponds to a potentially valuable
feature vector.  Given this representation, the random forest model
can be trained, in \textit{R} notation, to predict the labels via
$v_L \sim  \bm{F}$.
Our segmentation protocol involves training and application
of two RF models in succession 
(see Figure \ref{fig:pipeline}) to which we refer as ``stages.''  The basic idea is that
we generate a set of feature images used as input to the
first RF model (or first stage) which produces a voxelwise 
probabilistic tissue estimate.  More precisely, in the RF framework, each prediction sample (i.e. the feature vector at each voxel),
is propagated through each tree of the ensemble where it is labeled as belonging to a specific
class.  
These ``votes'' are converted to voxelwise probabilistic
estimates for each class via standard mechanisms used in random forest
models, e.g. \cite{liaw2002}.
The seven labels of brain/tumor tissue type denote:
\begin{itemize}
\item cerebrospinal fluid,
\item grey matter,
\item white matter,
\item edema, 
\item non-enhancing tumor (including low-grade tumor center), 
\item enhancing tumor (excluding necrotic center), and 
\item abnormal necrotic center or necrocyst in high-grade gliomas.
\end{itemize}
We then use these output tissue map estimates as spatial 
priors for generating a second set of geometrically refined image features.  These
are used as input for a second RF model application (or
second stage),
the output of which constitute the final tumor segmentation estimate.
Note that some feature images are included directly and
  without modification in both stages such as
the asymmetry features described in the previous section.
